---
title: "Appendix A"
date: "`r Sys.Date()`"
contact: claire.descombes@insel.ch
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
    theme: paper
    code_folding: show
    math_method: katex
subtitle: "R for physicians"
bibliography: C:/GitHub/rforphysicians/docs/Rforphysicians.bib
link-citations: yes
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {rmarkdown::render(input)})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE, 
  fig.height = 5,
  fig.width = 6,
  eval = TRUE
)
library(tidyverse)
```

| Author           |                                                                 |
|:-----------------|:----------------------------------------------------------------|
| Name             | Claire Descombes                                                |
| Affiliation      | Universitätsklinik für Neurochirurgie, Inselspital Bern         |
| Degree           | MSc Statistics and Data Science, University of Bern             |
| Contact          | [claire.descombes@insel.ch](mailto:claire.descombes@insel.ch)   |
The reference material for this course, as well as some useful literature to deepen your knowledge of R, can be found at the bottom of the page.

# NHANES data sets

The National Health and Nutrition Examination ([`NHANES`](https://www.cdc.gov/nchs/nhanes/)) Survey from 2011-2012 assessed overall health and nutrition of adults and children in the United States and was conducted by the National Center for Health Statistics (NCHS). The data sets can be found in the [`data_sets` folder](https://github.com/ClaireMargaux/rforphysicians/tree/main/data_sets).

| Dataset               | NHANES Code | Description                                                       | File                                    |
| --------------------- | ----------- | ----------------------------------------------------------------- | --------------------------------------- |
| Demographics          | DEMO\_G     | Age, sex, race/ethnicity, income, education                       | `DEMO_G.csv`                            |
| Blood Pressure        | BPX\_G      | Systolic/diastolic blood pressure, number of readings             | `BPX_G.csv`                             |
| Body Measures         | BMX\_G      | Height, weight, BMI, waist circumference                          | `BMX_G.csv`                             |
| Smoking Questionnaire | SMQ\_G      | Smoking habits                                                    | `SMQ_G.csv`                             |
| Mortality Data        | NDI         | Mortality follow-up data (status, follow-up time, cause of death) | `NHANES_2011_2012_MORT.dat`            |

💡 The codebook for each data set can be accessed either on the [`NCHS website`](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2011) or directly in `R` using the function `nhanesCodebook(nh_table, colname)` from the package `nhanesA` (which I used to download the data).

```{r}
library(nhanesA)

# An example here
nhanesCodebook('SMQ_G', 'SMQ020')
```

Follows a list of some of the key variables from the NHANES data sets.

## Demographics (DEMO_G)

| Variable Name | Description                                           |
| ------------- | ----------------------------------------------------- |
| RIDAGEYR      | Participant's age in years at screening               |
| RIAGENDR      | Participant's gender                                  |
| RIDRETH1      | Race/Hispanic origin                                  |
| DMDBORN4      | Country of birth                                      |
| DMDCITZN      | Citizenship status                                    |
| DMDMARTL      | Marital status                                        |
| DMDEDUC2      | Education level                                       |
| RIDEXPRG      | Pregnancy status at exam                              |
| INDHHIN2      | Annual household income                               |
| INDFMIN2      | Annual family income                                  |
| DMDHHSIZ      | Total number of people in the household               |
| DMDFMSIZ      | Total number of people in the family                  |
| DMDHHSZA      | Number of children aged 5 or younger in the household |
| DMDHRAGE      | Age of the household reference person                 |
| DMDHRMAR      | Marital status of the household reference person      |
| DMDHRGND      | Gender of the household reference person              |
| DMDHREDU      | Education level of the household reference person     |
| AIALANGA      | Language of the ACASI interview                       |

## Blood Pressure (BPX_G)

| Variable Name | Description                                        |
| ------------- | -------------------------------------------------- |
| BPXSY1        | Systolic blood pressure (first reading) in mm Hg   |
| BPXSY2        | Systolic blood pressure (second reading) in mm Hg  |
| BPXSY3        | Systolic blood pressure (third reading) in mm Hg   |
| BPXDI1        | Diastolic blood pressure (first reading) in mm Hg  |
| BPXDI2        | Diastolic blood pressure (second reading) in mm Hg |
| BPXDI3        | Diastolic blood pressure (third reading) in mm Hg  |
| BPXPLS        | 60-second pulse (30-second pulse multiplied by 2)  |
| BPXPULS       | Pulse regular or irregular?                        |
| BPAARM        | Arm selected                                       |
| BPQ150C       | Had coffee in the past 30 minutes?                 |
| BPQ150A       | Had food in the past 30 minutes?                   |
| BPQ150B       | Had alcohol in the past 30 minutes?                |
| BPQ150C       | Had coffee in the past 30 minutes?                 |
| BPQ150D       | Had cigarettes in the past 30 minutes?             |
| BPXCHR        | 60 sec HR (30 sec HR * 2)                          |

## Body Measures (BMX_G)

| Variable Name | Description                  |
| ------------- | ---------------------------- |
| BMXWT         | Weight (kg)                  |
| BMXHT         | Standing height (cm)         |
| BMXBMI        | Body Mass Index (kg/m²)      |
| BMXWAIST      | Waist circumference (cm)     |
| BMXARML       | Upper arm length (cm)        |
| BMXARMC       | Upper arm circumference (cm) |
| BMXLEG        | Upper leg length (cm)        |

## Smoking Questionnaire (SMQ_G)

| Variable Name | Description                                 |
| ------------- | ------------------------------------------- |
| SMQ020        | Smoked at least 100 cigarettes in life      |
| SMD030        | Age started smoking cigarettes regularly    |
| SMQ040        | Do you now smoke cigarettes?                |
| SMQ050Q       | How long since quit smoking cigarettes      |
| SMQ050U       | Unit of measure (day/week/month/year)       |
| SMD055        | Age last smoked cigarettes regularly        |
| SMD057        | # cigarettes smoked per day when quit       |
| SMQ077        | How soon after waking do you smoke          |
| SMD641        | # days smoked cigs during past 30 days      |
| SMD100BR      | Cigarette Brand/sub-brand                   |
| SMQ670        | Tried to quit smoking                       |
| SMQ621        | Cigarettes smoked in entire life            |

## Mortality data set (NHANES_2011_2012_MORT_2019_PUBLIC)

| Variable Name     | Description                                                                               |
| ----------------- | ----------------------------------------------------------------------------------------- |
| MORTSTAT      | Mortality status at the end of follow-up (0 = assumed alive, 1 = assumed deceased)        |
| PERMTH\_INT   | Person-months of follow-up from the NHANES interview date to date of death or censoring   |
| PERMTH\_EXM   | Person-months of follow-up from the NHANES examination date to date of death or censoring |
| UCOD\_LEADING | Underlying cause of death, grouped into leading cause categories (public-use)             |
| DIABETES      | Indicator if diabetes was listed as a cause of death (public-use)                         |
| HYPERTEN      | Indicator if hypertension was listed as a cause of death (public-use)                     |


\newpage

# References

---
nocite: '@*'
...